Towards a standardized evaluation of imputation methodology

### 2022-09-19 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
latest

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues with assessing imputation method performance. In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation. For example, focusing on one performance measure over another, may yield different conclusions about imputation efficacy. In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods. Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked. %(i.e. the complete-data model used to estimate the comparative truth). % The choice of performance measures might inadvertently influence simulation results.

@@

(ii) The estimates should have a proper coverage rate. Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}. Under-coverage indicates that the procedure is not confidence valid and may lead to invalid inference. Undercoverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both. Over-coverage, though technically not confidence invalid, would indicate that efficiency could still be gained. In such cases, a narrower interval exists for which the method would still yield confidence valid result.

>>>>

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues when assessing imputation method performance. In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation. For example, focusing on one performance measure over another, may yield different conclusions about imputation efficacy. In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods. Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked. %(i.e. the complete-data model used to estimate the comparative truth). % The choice of performance measures might inadvertently influence simulation results.

@@

(ii) The intervals for estimates should have a valid coverage of the population (i.e. true) value. Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}. Under-coverage indicates that the procedure is not confidence valid and may lead to invalid inference. Undercoverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both. Over-coverage, though technically not confidence invalid, would indicate that efficiency could still be gained. In such cases, a narrower interval exists for which the method would still yield confidence valid result.
